---
title: How to start a data blog with R + Hugo + Blogdown in 11 short steps
author: Glenn
date: '2017-12-15'
slug: how-to-start-a-data-blog-with-r-hugo-blogdown-in-11-short-steps
categories:
  - R tutorial
tags:
  - R
  - blogdown
  - hugo
---

I just finished creating this blog to share stories, teach R skills, create environemtnal data stories, and most importantly - shamelessly self-promote. It makes sense to share with you a minimal formula for starting your own data blog with R.  

Yihui Xie, the master behind Rmarkdown, has kindly created a graet framework for blogging with R - blogdown. Blogdown creates Hugo (one of the most popular open-source static site generators) websites, and allows you to write new content using markdown or Rmarkdown. This makes it really easy to create posts sharing R code and workflows. He has created a great, comprehensive guide available [here](https://bookdown.org/yihui/blogdown/). But the process s so simple you can achieve it in 11 simple steps:

1.Create a new project in Rstudio in a new folder  
  
2.Install the blogdown package
```{r eval=FALSE}
install.packages("blogdown")
```
  
3.Install Hugo (from within R)
```{r eval=FALSE}
blogdown::install_hugo()
```

4.Choose a theme for your blog from [https://themes.gohugo.io/](https://themes.gohugo.io/) - beware not all of them have been tested with blogdown  
  
5.Create your new site from this template. The string we use for the theme is the github site where the theme is hosted. The example below uses the hugo-academic theme hosted at [https://github.com/gcushen/hugo-academic](https://github.com/gcushen/hugo-academic)
```{r eval=FALSE}
blogdown::new_site(theme = "gcushen/hugo-academic")
```
  
6.This will automatically download all the files for your site into your project folder. You can now start editing these files to customise your site. 
 - Edit the `config.toml` file to change theme settings and general paramters of your site such as author, site description etc
 - Add images that you will use to the folder `static/images` in your project home directory
 - Edit your sites content by editing the pages within the `content` folder  in your project home directory. Here you can add new blog posts, or editing existing content
   
7. You can also add new blog posts using the handy Rstudio addin that comes with blogdown, just click `Tools > Addins > Browse Addins` and select `New Post`. You can specify the post title, dates, tags and categories using this tool.   
The same functionality can be acheived using a blogdown function:

```{r eval=FALSE}
blogdown::new_post()
```

8. Now you have a website, you probably want to share it with others. First create a repo on github for your website. Then initialize your project folder and push it to that repo. You will want to create a `.gitgnore` to prevent unessasary files from being pushed to git. An example `.gitignore` would be:

> public/  
> *.Rproj    
> *.Rhistory  
> .Rproj.user/

9. Create an account at [Netlify](http://www.netlify.com), it's free. Choose the option to create a new site from git, after loggining into your github account, select the repository containing your site.
